% I may change the way this is done in a future version, 
%  but given that some people needed it, if you need a different degree title 
%  (e.g. Master of Science, Master in Science, Master of Arts, etc)
%  uncomment the following 3 lines and set as appropriate (this *has* to be before \maketitle)
% \makeatletter
% \renewcommand {\@degree@string} {Master of Things}
% \makeatother

\title{Completeness Theorems for Behavioural Distances and Equivalences}
\author{Wojciech Krzysztof R\'{o}\.{z}owski}
\department{Department of Computer Science}

\maketitle
\makedeclaration

\begin{abstract} % 300 word limit
 In theoretical computer science it is customary to provide expression languages for representing the behaviour of transition systems and to study formal systems for reasoning about equivalence or similarity of behaviours represented by expressions of the interest. The key example of this approach are Kleene's regular expressions, a specification language for deterministic finite automata, as well as complete axiomatisations of language equivalence of regular expressions due to Salomaa and Kozen.

The first part of this thesis studies axiomatisations of behavioural distances. Originally considered for probabilistic and stochastic systems, behavioural distances provide a quantitative measure of the dissimilarity of behaviours that can be defined meaningfully for a variety of transition systems. As a first contribution, we consider deterministic automata and we provide a sound and complete quantitative inference system for reasoning about a shortest-distinguishing-word distance between languages represented by regular expressions. Then, we move on to more complicated case of behavioural distance of Milner's charts, which provide a compelling setting for studying behavioural distances because they shift the focus from language equivalence to bisimilarity. As a syntax of choice, we rely on string diagrams, which provide a rigorous formalism that enables compositional reasoning by supporting a variable-free representation where recursion naturally decomposes into simpler components.

The second part focuses on generative probabilistic transition systems and presents a sound and complete axiomatisation of language equivalence of behaviours specified through the syntax of probabilistic regular expressions (PRE), a probabilistic analogue of regular expressions denoting probabilistic languages in which every word is assigned a probability of being generated. The completeness proof makes use of technical tools from the recently developed theory of proper functors and convex algebra, arising from the rich structure of probabilistic languages. 


\end{abstract}

\begin{impactstatement}
\textbf{Outside academia:} The discipline of formal verification enables making precise logical statements about computer systems to guarantee their correctness. This relies on the study of models of computation, which are mathematical objects representing the semantics of systems of interest. In theoretical computer science, it is customary to model computations as transition systems. This thesis is part of a larger research programme aimed at providing specification languages. for representing transition systems and studying formal systems for reasoning about the equivalence or similarity of systems represented by the syntax. One of the central kind of problems attached to this field of research are completeness problems, that concern showing that every semantic equivalence or similarity of transition systems can be witnessed through the means of axiomatic manipulation. Having a complete axiomatisation allows one to fully resort to syntactic reasoning, which is particularly amenable to implementation and thus desirable from the automated reasoning point of view. 

Kleene algebra (KA)~\cite{Kozen:1994:Completeness} is a central example of such specification language. KA is a foundation of tools relevant to industry, such as NetKAT~\cite{Anderson:2014:NetKAT}, which enables reasoning about the behaviour of packet-passing Software-Defined Networks, or cf-GKAT~\cite{Zhang:2025:CFGKAT}, which can be used to certify the correctness of decompilation algorithms. This thesis particularly focuses on the semantic notions of behavioural distances and probabilistic language equivalence, where it makes its main contributions. Behavioural distances~\cite{Breugel:2001:Towards,Desharnais:2004:Metrics} replace the strict notion of equivalence of states with a more liberal quantitative measure of dissimilarity. This is particularly desirable for stochastic or probabilistic systems, where a tiny observed perturbation would lead to inequivalence of states. Behavioural distances have been successfully applied to Markov decision processes within reinforcement learning (RL), with the key example being MICo (matching under independent couplings) distance~\cite{Castro:2021:MICo}. Additionally, probabilistic language equivalence, axiomatised in Chapter 4 of this thesis, underpins Apex~\cite{Kiefer:2012:APEX}, an automated equivalence checker for probabilistic programs.

\textbf{Inside academia:} This thesis makes original contributions within the field of theoretical computer science. The results in \Cref{chapter2} are an adaptation of a concrete completeness result for behavioural distance of probabilistic transition systems to an instance of an abstract coalgebraic framework. Besides the basic examples provided recently by Lobbia et al~\cite{Lobbia:2024:Quantitative}, the content of \Cref{chapter3} is the first work to propose a complete axiomatisation of a quantitative calculus of string diagrams through a systematic axiomatic foundation. Finally, \Cref{chapter4} provides an alternative axiomatisation of language equivalence of generative probabilistic transition systems~\cite{Glabbeek:1995:Reactive} through a simple generalisation of Kleene’s regular expressions. The completeness result makes use of recently developed theory of proper functors~\cite{Milius:2018:Proper} and provides further evidence that the use of coalgebras for proper functors provides a good abstraction for completeness theorems.

The entire material presented in this thesis has been accepted or is under review for several highly-ranked conferences in theoretical computer science. The results of this thesis have been disseminated to the computer science community through a series of seminar talks across the United Kingdom, the United States, and Germany. 
\end{impactstatement}


\begin{acknowledgements}
First and foremost, I would like to thank my supervisor, Alexandra Silva, for giving me the opportunity to embark on this PhD and for her unwavering support over the past four years. I am deeply grateful for the many fantastic opportunities she exposed me to, as well as for her invaluable academic and career advice. I sincerely appreciate her time, energy, and trust in me. I am also thankful to Alexandra for hosting my visits to Cornell and for making my time in Ithaca both enjoyable and productive.

I am very grateful to my examiners, Filippo Bonchi and Clemens Kupke, for reading my thesis so thoroughly and offering valuable suggestions to improve the final version of the manuscript. I would also like to thank Samson Abramsky and Robin Hirsch, who examined my first- and second-year vivas. Thanks to all of them, every examination I encountered during my PhD was a genuinely positive experience.

I would like to acknowledge all of my co-authors on the papers I worked on during my PhD. Thank you to Alexandra, Barbara, Dexter, Fabio, Gabriele, Johanna, Jurriaan, Keri, Matina, Paul, Ralph, Sebastian, Spencer, Stefan, Tobias, and Todd. I am especially grateful to my dear friend Todd, who was my first collaborator and taught me so much during the early days of my PhD.

At the midpoint of my PhD, I had the privilege of participating in the Adjoint School 2023 in Applied Category Theory. Collaborating with Barbara König was one of the highlights of my PhD and introduced me to behavioural distances, which ultimately became the central focus of my thesis. I am very grateful to Barbara for hosting two highly productive visits to the University of Duisburg-Essen.

I would also like to thank Giorgio Bacci and Ana Sokolova for the insightful discussions that contributed to the content of Chapters 2 and 4 of this thesis, respectively.

In the summers of 2023 and 2024, I had the privilege of interning with the Automated Reasoning Group at Amazon Web Services in Seattle, where I worked alongside many wonderful people. I am especially grateful to my managers, George-Axel Jaloyan and Daniel Schoepe, as well as Sean McLaughlin, for making both internships such a positive and rewarding experience. I would also like to thank Rustan Leino and Leo de Moura for the many valuable discussions and conversations we shared during my time at Amazon. A special thanks goes to my friend Tabea, who was interning at the same time and became my companion in exploring Seattle and the Pacific Northwest. I am also very thankful to George for organising several fun activities that made my stay in Seattle even more enjoyable.

I would like to thank the organisers of the Dagstuhl Seminar on Quantitative Logic and Behavioural Metrics, as well as the two Bellairs Workshops on Probabilistic Semantics and Quantitative Logic, for inviting me to participate. As a junior academic, it was a real pleasure to engage with the research community in my field in such stimulating and welcoming environments.

I would not have embarked on a PhD if it weren’t for the inspiring academics at the University of Southampton, who sparked my passion for theoretical computer science and supported me throughout my academic journey. I am especially grateful to Julian Rathke for supervising my Bachelor's thesis and for his invaluable help during the PhD application process. My heartfelt thanks also go to my exceptional lecturers, Corina Cîrstea and Paweł Sobociński, whose teaching made me fall in love with theory and who have supported me ever since. I would also like to thank Enrico Gerding, who was my personal academic tutor at the time, for his guidance and support.

If I were to properly thank each of my friends, I would probably need to write a separate thesis. During my time at UCL, I was lucky to be surrounded by many incredible people - close friends, flatmates, concert and rave companions, who made those years full of joy, laughter, and unforgettable experiences. Thank you Alesandro, Antonio, Billy, Cheng, Delia, Fred, Gerco, Kevin, Leo, Linpeng, Louis, Mateo, Jaš, Katya, Tiago, Rafał, Ralph, Robin, Sam, Stefan, and Yll.
Beyond the friends I made through work, I was also fortunate to have a wonderful support network of people who made the past four years not only bearable but truly fun, and who provided many much-needed positive distractions. Thank you Aiden, Behrad, Bennie, Daniela, Dominik, Gaby, Iwo, Jackie, Juliusz, Lucy, Maciek, Molly, Natalie, Nima, Ola, Olga, Patryk, and Zuza.

I would like to thank my parents, to whom I dedicate this thesis. I am deeply grateful for everything they have done for me throughout my life, and I cannot overstate how much their love and support have meant to me. Simply put, I would never have made it this far without them. I feel incredibly lucky to call this pair of amazing human beings my parents. Additionally, I am deeply grateful to my aunt Grażyna for her unwavering financial support during my undergraduate years, which greatly helped me along the way.

Finally, I would like to thank my wonderful girlfriend, Kinga, for her unwavering love and support. I feel incredibly lucky to have such a caring and inspiring partner by my side. The time I have spent with her has been one of the greatest highlights of my journey, and I am truly honoured to have her in my life.
\end{acknowledgements}

\setcounter{tocdepth}{2} 
% Setting this higher means you get contents entries for
%  more minor section headers.

\tableofcontents
\listoffigures
%\listoftables

